Implementing \Kwd{DNS} in \Kwd{C0}

### 2017-12-08 paper.tex
add a bunch of text to the paper

\documentclass{article}
\usepackage[tt=false]{libertine}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage[colorlinks=true]{hyperref}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{microtype}

@@

\section{Background and Related Work}

\Kwd{Domain Name System} (\Kwd{DNS}) is a hierarchical database that
maps domain names to IP addresses in a distributed manner augmented
with caching to achieve a better
performance~\cite{rfc:1034,rfc:1035,mun-lee:2005}. In fact, there are
three major parts:

@@

\subsection{C0}\label{sec:c0}

The \Kwd{C0} programming language is a restricted subset of \Kwd{C}
that provides implicit memory safety to
programmers~\cite{pfenning:c0-reference}. Restrictions on pointer
arithmetic and casting, for example, allow the language to detect and
prevent any attempt to access an array out of its bounds. Furthermore,
there is no need for error-prone manual memory management, since
\Kwd{C0} is itself garbage-collected.

@@

% \Kwd{Concurrent C0} (\Kwd{CC0}) is an extension of \Kwd{C0} that
% offers concurrency based on message passing by providing processes as
% the units of concurrent execution and channels as their message
% buffers~\cite{willsey-prabhu-pfenning:2016}. Concurrent processes
% communicate over channels using a protocol that is disciplined by
% session types~\cite{caires-pfenning-toninho:2013,
%   balzer-pfenning:2017}. In this protocol, each channel is typed such
% that it can either send or receive a message and, thus, it
% deliberately models the client/provider relation.

@@

\subsection{Related Work}

\paragraph{Nail} TODO (Jon) evaluate Nail.
\nocite{bangert:2014}

\paragraph{The \Kwd{Fox} Project}
An earlier example of an implementation of the DNS protocol in a safe
and formally specified language is given by the \Kwd{Fox} project,
undertaken at Carnegie Mellon University in the
1990s~\cite{biagioni-harper-lee-milnes:1994,
  biagioni-harper-lee:2001}. Their principal artifact was a library
called \Kwd{FoxNet} which implemented an array of low-level network
infrastructure, including \Kwd{TCP}, \Kwd{IP}, \Kwd{ARP} and \Kwd{DNS}
in the \Kwd{Standard ML} language.

@@

\paragraph{Other implementations} More recently, safe languages like
\Kwd{Rust} and \Kwd{Go} have been gaining popularity. Both languages
have hosted implementations of the \Kwd{DNS}
protocol~\cite{github:trust-dns,github:miekg-dns}

@@

typedef struct header* header_t;
  \end{lstlisting}

@@

typedef struct domain_name* domain_name_t;

struct resource_record {
  domain_name_t name;
  int type;
  int class;
  int ttl;
  int rd_length;
  int[] rdata;
};

typedef struct resource_record* resource_record_t;
  \end{lstlisting}
  \caption{Example DNS data structures in our \Kwd{C0} implementation.}\label{fig:c0-data-structures}
\end{figure}

@@

\begin{lstlisting}
int parse_domain_name(int cursor, int[] msg, domain_name_t* dest);
int parse_resource_record(int cursor, int[] msg, resource_record_t dest);
int parse_header(int cursor, int[] msg, header_t dest);
\end{lstlisting}

These functions take as an argument a cursor (an index in bits into
the message), an array of 32-bit words (the entire input message), and
pointer to a structure into which to write the parsed data; they
return a new cursor. It is important for the modularity of our
development that cursors represent not indices into the 32-bit cells
of the message, but rather bit-indices; this is because specific parts
of DNS data structures may only be a few bits long, so their parsers
cannot move the cursor forward a full word's length.
%
A sample of our parsing code can be found in the supplementary
appendix to this proposal (\S\ref{appendix:decompression}). Of
course, we will also implement the inverses to the above functions.

@@

% For performance analysis, we plan to evaluate our nameserver for
% latency and throughput, as these appear to be the most important
% metrics for \Kwd{DNS} performance. We plan to do this using one of two
% pieces of software\cite{github:dnsperf1, github:dnsperf2}
% (confusingly, both named \texttt{dnsperf}), which read queries from a
% file and then send these requests to a nameserver. The tools mentioned
% in~\cite{github:dnsperf2} seem better-suited to our needs, as those
% of~\cite{github:dnsperf1} appear to not give latency information in a
% useful fashion.

@@

\bibliographystyle{abbrv}
\bibliography{bibtex-references/refs,project}

@@

    if (len >= 0xc0) {
        // case: pointer
        // read the last 14 bits of this byte as a
        // pointer into the message (in octets)
        cursor_set(cur, orig_ix + 2);
        int ptr = read_bits(cur, data, 14);

>>>>

\documentclass{article}
\usepackage[tt=false]{libertine}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage[colorlinks=true, citecolor=red]{hyperref}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{microtype}
\usepackage{natbib}

@@

\section{Background}

\Kwd{Domain Name System} (\Kwd{DNS}) is a hierarchical database that
maps domain names to IP addresses in a distributed manner augmented
with caching to achieve a better
performance~\citep{rfc:1034,rfc:1035,mun-lee:2005}. In fact, there are
three major parts:

@@

\subsection*{The \Kwd{C0} Language}\label{sec:c0}

The \Kwd{C0} programming language is a restricted subset of \Kwd{C}
that provides implicit memory safety to
programmers~\citep{pfenning:c0-reference}. Restrictions on pointer
arithmetic and casting, for example, allow the language to detect and
prevent any attempt to access an array out of its bounds. Furthermore,
there is no need for error-prone manual memory management, since
\Kwd{C0} is itself garbage-collected.

@@

% \Kwd{Concurrent C0} (\Kwd{CC0}) is an extension of \Kwd{C0} that
% offers concurrency based on message passing by providing processes as
% the units of concurrent execution and channels as their message
% buffers~\citep{willsey-prabhu-pfenning:2016}. Concurrent processes
% communicate over channels using a protocol that is disciplined by
% session types~\citep{caires-pfenning-toninho:2013,
%   balzer-pfenning:2017}. In this protocol, each channel is typed such
% that it can either send or receive a message and, thus, it
% deliberately models the client/provider relation.

@@

\section{Related Work}

\subsection*{Nail: parsing and generating data formats}

Marshalling and unmarshalling low-level data formats
is a highly error-prone process, requiring a great degree of care for
things like byte alignment, sign, etc.; mistakes in this kind of code
are responsible for numerous security vulnerabilities, especially when
written in unsafe languages like \Kwd{C}.

To ameliorate this difficulty, \citeauthor{bangert:2014} have
introduced \Kwd{Nail}, a specialized language for specifying low-level
data formats. \Kwd{Nail} specifications can then be compiled into real
parsers and formatters; as part of their evaluation, the authors of
\Kwd{Nail} have used it to develop a ``authoritative DNS server''.

\Kwd{Nail} has a number of interesting and extremely useful features,
including \emph{dependent fields}: these are used when the sort of
data in one field depends on the actual data of another field: for
instance, when a \verb|length| field is followed by an array of bytes
of that length.
%
Additionally, \Kwd{Nail} supports injecting user-supplied pairs of
partial inverses to wrap data in the generated parser and formatter;
this is crucial for compressed or encrypted data.

\paragraph{Evaluation}

We regard the \Kwd{Nail} project as an interesting point in the design
space, enabling rapid development of tools that interact with
low-level data formats. As far as safety is concerned, it is necessary
to trust two things:

\begin{enumerate}
\item The internal implementation of \Kwd{Nail}, which is written in
  \Kwd{C}. It is not clear to us what it means to trust \Kwd{C} code,
  because the ``semantics'' of \Kwd{C} code is completely informal and
  indefinite.
\item The user-provided implementations of any functions which are
  used for compression, decompression, etc. Indeed, there is reason to
  doubt the correctness of such code, even when written by experts;
  for instance, the \Kwd{DNS} demo in the \Kwd{Nail} development
  contains an operator precedence bug~\citep{nigeltao:ticket}.
\end{enumerate}

The issue described above suggests that while \Kwd{Nail} addresses
many difficulties, the most error-prone aspects of protocol
implementation are out of its scope. We think it would be an
interesting and worthwhile project to pursue a \emph{verified} version
of \Kwd{Nail} as a library for \Kwd{Coq}~\citep{coq:reference-manual},
enabling users to verify the correctness of their compression and
decompression functions relative to a formal approximation of
\Kwd{C}'s semantics.

Contrary to \Kwd{Nail}, our approach was to code parsers and
formatters for the \Kwd{DNS} data structures \emph{manually}, but in
memory-safe and type-safe language. This means that while functional
correctness is not guaranteed, \emph{safety} is guaranteed.

\subsection*{The \Kwd{Fox} Project}
An earlier example of an implementation of the DNS protocol in a safe
and formally specified language is given by the \Kwd{Fox} project,
undertaken at Carnegie Mellon University in the
1990s~\citep{biagioni-harper-lee-milnes:1994,
  biagioni-harper-lee:2001}. Their principal artifact was a library
called \Kwd{FoxNet} which implemented an array of low-level network
infrastructure, including \Kwd{TCP}, \Kwd{IP}, \Kwd{ARP} and \Kwd{DNS}
in the \Kwd{Standard ML} language.

@@

\subsection*{Other implementations}

More recently, safe languages
like \Kwd{Rust} and \Kwd{Go} have been gaining popularity. Both
languages have hosted implementations of the \Kwd{DNS}
protocol~\citep{github:trust-dns,github:miekg-dns}

@@

typedef struct header* header;
  \end{lstlisting}

@@

typedef struct domain_name* domain_name;

struct resource_record {
  domain_name name;
  int type;
  int class;
  int ttl;
  int rd_length;
  int[] rdata;
};

typedef struct resource_record* resource_record;
  \end{lstlisting}
  \caption{Example DNS data structures in our \Kwd{C0} implementation.}\label{fig:c0-data-structures}
\end{figure}

@@

\begin{lstlisting}
domain_name parse_domain_name(cursor cur, int[] data);
resource_record parse_resource_record(cursor cur, int[] data);
header parse_header(cursor cur, int[] data);
\end{lstlisting}

\paragraph{The cursor abstraction}
In order to enable a style of coding which is less error-prone than
manual cursor-passing, we implemented an abstract type of imperative
cursors, which can be advanced or retreated by a certain number of
bits. Then, our low-level bit arithmetic utilities can be used to read
(or write) a certain number of bits at a certain offset from (resp.\
to) our stream of words.

It is important for the modularity of our development that cursors
represent not indices into the 32-bit cells of the message, but rather
bit-indices; this is because specific parts of DNS data structures may
only be a few bits long, so their parsers cannot move the cursor
forward a full word's length.

A sample of our parsing code can be found in the supplementary
appendix to this proposal (\S\ref{appendix:decompression}).

\paragraph{The implemented fragment}

We have implemented marshalling and unmarshalling for a sizable
fragment of the \Kwd{DNS} protocol, omitting only some rare or
experimental resource record types. Our parser for domain names
supports compression as specified in~\cite{rfc:1035} (see
Appendix~\ref{appendix:decompression} for our domain name routine);
however, during formatting we do not attempt to re-compress domain
names. This would be necessary for a serious implementation.

\paragraph{Tests}

We have implemented several tests to ensure correct roundtrip behavior
of the marshalling and unmarshalling routines by hand-coding instances
of the \Kwd{DNS} message structure.

\section{Evaluation}

\subsection{Correctness and Safety}

By virtue of having written the majority of our \Kwd{DNS} server in
the memory- and type-safe \Kwd{C0} language, we enjoy certain safety
guarantees that are either impossible or intractable to achieve for
mainstream \Kwd{C}-based implementations. In return, we have sacrified
some performance, but it is hard to tell how much of our performance
trade-off was due to garbage collection vs.\ the unoptimized character of
our implementation (but see Section~\ref{sec:performance}).

\paragraph{Contracts in \Kwd{C0}}

One feature of \Kwd{C0} that we have found completely invaluable is
its notion of \emph{contract}, which are a lightweight form of
assertion which wraps the input and output boundaries of
procedures. Consider for instance the following code from our
low-level bit manipulation library:

\begin{lstlisting}
int get_mask(int index, int size)
//@requires size < 32 && size >= 0;
//@requires index < 32 && index >= 0;
{
    return ((1 << size) - 1) << index;
}

int get_bits(int index, int size, int data)
//@requires size < 32 && size >= 0;
//@requires index < 32 && index >= 0;
{
    return (data & get_mask(index, size)) >> index;
}

int set_bits(int data, int index, int size, int value)
//@requires size < 32 && size >= 0;
//@requires index < 32 && index >= 0;
{
    return (data & ~(get_mask(index, size))) | (value << index);
}
\end{lstlisting}

Here, we used contracts to ensure that we would not suffer from bit
shift range errors. The code did not begin with these contracts, but
in the course of debugging, these contracts were introduced; in every
instance, introducing contracts (shifting blame closer and closer to
call-sites) enabled us to easily determine and liquidate bugs.

The \Kwd{C0} compiler can be instructed to check contracts dynamically
during execution, or to generate code which ignores them. This is
helpful, since it is important to have debuggable errors when
developing, but in production one may not wish to incur the cost of
dynamic contract checks. While dynamic execution of contract checks
is already very useful, we feel that more advanced techniques such as
abstract interpretation and its special case, symbolic execution,
would be even more useful.

\subsection{Performance}\label{sec:performance}

\textit{Klaas, please write this.}

@@

% For performance analysis, we plan to evaluate our nameserver for
% latency and throughput, as these appear to be the most important
% metrics for \Kwd{DNS} performance. We plan to do this using one of two
% pieces of software\citep{github:dnsperf1, github:dnsperf2}
% (confusingly, both named \texttt{dnsperf}), which read queries from a
% file and then send these requests to a nameserver. The tools mentioned
% in~\cite{github:dnsperf2} seem better-suited to our needs, as those
% of~\cite{github:dnsperf1} appear to not give latency information in a
% useful fashion.

@@

\bibliographystyle{abbrvnat}
\bibliography{bibtex-references/refs,project}

@@

    if (len >= 0xc0) {
        // case: pointer
        // read the last 14 bits of this word as a
        // pointer into the message (in octets)
        cursor_set(cur, orig_ix + 2);
        int ptr = read_bits(cur, data, 14);
